The Space of Feasible Execution Times for Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-09-27 dit-paper_ieee.tex

		This system is equivalent to multiple instances (one per combination of $a_i$
		values) of the Chinese Remainder Problem (CRP) where the goal is to find the
		value of $t_d$ such that it is congruent to $a_i+O_i$ modulo $T_i$ for all
		tasks $\tau_i$ in $\tau$. To find the FPDIT it is to the best of our knowledge
		necessary to solve the problem on all combinations of $a_i$ values and then
		to compute the minimum solution. We define $x_i = a_i+O_i$ and $n_i = T_i$ for
		clarity in the context of the CRP. One instance of this problem is guaranteed
		to have at least one solution modulo $H$ for given values of $a_i$ if and only if
		Condition~\ref{eq:CRPcond} is satisfied \cite{knuth1969art}:
		\begin{equation}
			\label{eq:CRPcond}
			x_i \equiv x_j \; (\operatorname{mod} \; \operatorname{gcd}(n_i, n_j))
			\; \forall \tau_i,\tau_j \in \tau
		\end{equation}
		Condition~\ref{eq:CRPcond} means that DITs might still exist even though in
		the general case task periods will not necessarily be pairwise coprime.
% 		We now present an example of such a system.\\
%
% 		\begin{center}
% 			\begin{tabular}{|r|c|c|c|c|}
% 				\hline
% 							& $O_i$ & $C_i$ & $D_i$ & $T_i$ \\ \hline
% 				$\tau_1$ 	& 0 	& $C_1$ & 3 	& 4\\ \hline
% 				$\tau_2$ 	& 2 	& $C_2$ & 7 	& 8\\ \hline
% 			\end{tabular}
% 		\end{center}
%
% 		Here, the possible values of $a_1$ are 3 and 4, and the possible values
% 		of $a_2$ are 7 and 8. For a FPDIT to exist, one of the following
% 		conditions must be fulfilled:
% 		\[
% 			\left[
% 			\begin{array}{c}
% 				3 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				3 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				4 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				4 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
% 			\end{array}
% 			\right.
% 		\]
% 		However neither is, which means the system has
% 		no FPDIT (and no other DIT than $t=0$).

	\subsection{Solving the modular system}
		A sketch of the construction algorithm to compute $t_d$ can be found in
		Algorithm~\ref{alg:algoCRP} (note that $x^{-1} \; (mod \; y)$ means the
		modular inverse of $x$ in the naturals modulo $y$). The congruences must
		first be modified in order to obtain pairwise coprime moduli. Each $n_i$
		is factored into prime powers $p^{k_1}_1 p^{k_2}_2 \cdots p^{k_m}_m = n_i$.
		Each original constraint $t_d \equiv x_i \; (mod \; n_i)$ is replaced by the
		equivalent set
		\begin{equation*}
			\begin{split}
				t_d \equiv x_i 	& \; (mod \; p^{k_1}_1) \\
				t_d \equiv x_i 	& \; (mod \; p^{k_2}_2) \\
								& \vdots \\
				t_d \equiv x_i 	& \; (mod \; p^{k_m}_m) \\
			\end{split}
		\end{equation*}
		This operation might create duplicate constraints, which are removed. As shown
		in \cite{knuth1969art}, the resulting system can be solved using
		Algorithm~\ref{alg:algoCRP}.

		\begin{algorithm}
			\caption{Gauss's CRP Algorithm}
			\label{alg:algoCRP}
			\begin{algorithmic}[1]
				\REQUIRE $H$ is the hyperperiod of the system
				\FORALL{$\tau_i \in \tau$}
					\STATE $Mj \leftarrow \frac{H}{n_i}$
					\STATE $invMj \leftarrow Mj^{-1} \; (mod \; n_i)$
					\STATE $e_i \leftarrow Mj \cdot invMj$
				\ENDFOR
				\RETURN $\sum\limits_{\tau_i \in \tau}{x_i \cdot e_i} \; (mod \; H)$
			\end{algorithmic}
		\end{algorithm}

@@

		Algorithm~\ref{alg:algoCRP} finds one DIT given one combination of $a_i$
		values. However using it in practice is very inefficient as there are
		exponentially many different combinations of $a_i$ values in the number of
		tasks, and resolving exponentially many CRP instances is prohibitively slow.
		A faster method for finding the FPDIT consists in merging all these CRP
		instances into one problem (finding all DITs), which can be solved faster by
		Algorithm~\ref{alg:multiACRP}, as a significant part of the intermediate
		results are shared between instances.

		%Review 3 : add formal complexity assessment of this algorithm

		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
				\REQUIRE $\forall \tau_i \in \tau: X_i$ is the set of possible $a_i+O_i
				\; \forall a_i \in [D_i,T_i]$ values for $\tau_i$
				\REQUIRE $\forall \tau_i \in \tau : n_i = T_i$
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE Modify $n_i$ and the content of $X_i$ by prime
				\\ decomposition to get pairwise coprime $n_i$ values

				\STATE $sumchunks \leftarrow \text{empty list}$
				\FORALL{$\tau_i \in \tau$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mj \leftarrow \frac{H}{n_i}$
					\STATE $invMj \leftarrow Mj^{-1} (mod \; n_i)$
					\FORALL{$x_i \in X_i$}
						\STATE append $(x_i \cdot Mj \cdot invMj)$ to $taskTerms$
					\ENDFOR
					\STATE $sumchunks[i] \leftarrow taskTerms$
				\ENDFOR

				\STATE $results \leftarrow \text{empty list}$
				\STATE append $0$ to $results$
				\FORALL{$\tau_i \in \tau$}
					\STATE $newresults \leftarrow \text{empty list}$
					\FORALL{$chunk \in sumchunks[i]$}
						\FORALL{$r \in results$}
							\STATE $\text{append} \; (r + chunk) \; \text{to} \; newresults$
						\ENDFOR
					\ENDFOR
					\STATE $results \leftarrow newresults$
				\ENDFOR
				\RETURN $r \; (mod \; H)$ for all $r \in results$
			\end{algorithmic}
		\end{algorithm}

		Both Algorithm~\ref{alg:algoCRP} and Algorithm~\ref{alg:multiACRP} require the
		computation of one modular inversion per $a_i$ value, which is done in a
		linear number of modular inversions per task. However, to
		find all DITs Algorithm~\ref{alg:algoCRP} has to be executed once for each possible
		combination of $a_i$ values, which is exponential in the number of tasks.
		Algorithm~\ref{alg:multiACRP} has the same theoretical complexity as the
		computation of all DITs still requires an exponential number of answers in the
		number of tasks, but it is nonetheless significantly faster in practice. The
		running time of Algorithm~\ref{alg:multiACRP} is also less sensitive to the
		number of different $a_i$ values per task, which means that this method
		remains practical when tasks deadlines are much smaller than periods.

>>>>

		This system is equivalent to multiple instances (one per combination of $a_i$
		values) of the Chinese Remainder Problem (CRP) where the goal is to find the
		value of $t_d$ such that it is congruent to $a_i+O_i$ modulo $T_i$ for all
		tasks $\tau_i$ in $\tau$. To find the FPDIT it is to the best of our knowledge
		necessary to solve the problem on all combinations of $a_i$ values and then
		to compute the minimum solution. We define $A_i = [D_i,T_i]$, $X_i = \{x_i
		= a_i+O_i \mid a_i \in A_i\}$ and $n_i = T_i$ for clarity in the context of
		the CRP.
		One instance of this problem is guaranteed to have at least one solution modulo $H$ for given values of $a_i$ if and only if
		Condition~\ref{eq:CRPcond} is satisfied \cite{knuth1969art}:
		\begin{equation}
			\label{eq:CRPcond}
			x_i \equiv x_j \; (\operatorname{mod} \; \operatorname{gcd}(n_i, n_j))
			\; \forall \tau_i,\tau_j \in \tau
		\end{equation}
		Condition~\ref{eq:CRPcond} means that DITs might still exist even though in
		the general case task periods will not necessarily be pairwise coprime.
% 		We now present an example of such a system.\\
%
% 		\begin{center}
% 			\begin{tabular}{|r|c|c|c|c|}
% 				\hline
% 							& $O_i$ & $C_i$ & $D_i$ & $T_i$ \\ \hline
% 				$\tau_1$ 	& 0 	& $C_1$ & 3 	& 4\\ \hline
% 				$\tau_2$ 	& 2 	& $C_2$ & 7 	& 8\\ \hline
% 			\end{tabular}
% 		\end{center}
%
% 		Here, the possible values of $a_1$ are 3 and 4, and the possible values
% 		of $a_2$ are 7 and 8. For a FPDIT to exist, one of the following
% 		conditions must be fulfilled:
% 		\[
% 			\left[
% 			\begin{array}{c}
% 				3 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				3 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				4 + 0 \equiv 7 + 2 \; (\operatorname{mod} \; 4 ) \\
% 				4 + 0 \equiv 8 + 2 \; (\operatorname{mod} \; 4 ) \\
% 			\end{array}
% 			\right.
% 		\]
% 		However neither is, which means the system has
% 		no FPDIT (and no other DIT than $t=0$).

	\subsection{Solving the modular system}
		A sketch of the construction algorithm to compute $t_d$ can be found in
		Algorithm~\ref{alg:algoCRP} (note that $x^{-1} \; (mod \; y)$ means the
		modular inverse of $x$ in the naturals modulo $y$). The congruences must
		first be modified in order to obtain pairwise coprime moduli. Each $n_i$
		is factored into prime powers $p^{k_1}_1 p^{k_2}_2 \cdots p^{k_m}_m = n_i$
		with $k_j$ the highest integer s.t. $p^{k_j}_j$ divides $n_i \; \forall j
		\; \mid 1 \leqslant j \leqslant m$. Each original constraint $t_d \equiv x_i
		\; (mod \; n_i)$ is replaced by the equivalent set
		\begin{equation*}
			\begin{split}
				t_d \equiv x_i 	& \; (mod \; p^{k_1}_1) \\
				t_d \equiv x_i 	& \; (mod \; p^{k_2}_2) \\
								& \vdots \\
				t_d \equiv x_i 	& \; (mod \; p^{k_m}_m) \\
			\end{split}
		\end{equation*}
		This operation (later called constraint \emph{factorisation}) might create
		duplicate constraints, which are removed.
		As shown in \cite{knuth1969art}, the resulting system can be solved using
		Algorithm~\ref{alg:algoCRP}. For ease of reading, we call $C$ the set of
		resulting constraints $c$ each written $t_d \equiv x_c \; (mod \;
		p_c)$.

		\begin{algorithm}
			\caption{Gauss's CRP Algorithm}
			\label{alg:algoCRP}
			\begin{algorithmic}[1]
				\REQUIRE $H$ is the hyperperiod of the system
				\FORALL{$c \in C$}
					\STATE $Mc \leftarrow \frac{H}{p_c}$
					\STATE $invMc \leftarrow Mc^{-1} \; (mod \; p_c)$
					\STATE $e_c \leftarrow Mc \cdot invMc$
				\ENDFOR
				\RETURN $\sum\limits_{c \in C}{x_c \cdot e_c} \; (mod \; H)$
			\end{algorithmic}
		\end{algorithm}

@@

		Algorithm~\ref{alg:algoCRP} finds one DIT given one combination of $a_i$
		values. However using it in practice is very inefficient as there are
		exponentially many different combinations of $a_i$ values in the number of
		tasks (the cartesian product of all the $A_i$ sets), and resolving
		exponentially many CRP instances is prohibitively slow.
		A faster method for finding the FPDIT consists in merging all these CRP
		instances into one problem (finding all DITs), which can be solved faster by
		Algorithm~\ref{alg:multiACRP}, as a significant part of the intermediate
		results are shared between instances. The first part consisting in splitting
		all constraints into sets of constraints with prime power moduli is slightly
		different. We define for each task $\tau_i$ and $a_j \in A_i$ :
		$C_{ij}$ is the set of constraints resulting from the factorisation of $t_d
		\equiv a_j+O_i \; (mod \; n_i)$ of the form $t_d \equiv x_c
		\; (mod \; p_c)$ if named $c$. We define $C = \cup\limits_{\tau_i \in \tau,
		a_j \in A_i}{C_{ij}}$, $P$ the set of all $p_c$ values appearing in
		constraints of $C$ and $C_{p}$ the subset of $C$ containing all constraints
		$c$ with $p_c = p$.

		\begin{algorithm}
			\caption{Modified Gauss Algorithm}
			\label{alg:multiACRP}
			\begin{algorithmic}[1]
				\REQUIRE $H$ is the hyperperiod of the system

				\STATE $sumchunks \leftarrow \text{empty map}$
				\FORALL{$p \in P$}
					\STATE $taskTerms \leftarrow \text{empty list}$
					\STATE $Mp \leftarrow \frac{H}{p}$
					\STATE $invMp \leftarrow Mp^{-1} (mod \; p)$
					\FORALL{$c \in C_p$}
						\STATE append $(x_c \cdot Mp \cdot invMp)$ to $taskTerms$
					\ENDFOR
					\STATE $sumchunks[p] \leftarrow taskTerms$
				\ENDFOR

				\STATE $results \leftarrow \text{empty list}$
				\STATE append $0$ to $results$
				\FORALL{$p \in P$}
					\STATE $newresults \leftarrow \text{empty list}$
					\FORALL{$chunk \in sumchunks[p]$}
						\FORALL{$r \in results$}
							\STATE $\text{append} \; (r + chunk) \; \text{to} \; newresults$
						\ENDFOR
					\ENDFOR
					\STATE $results \leftarrow newresults$
				\ENDFOR
				\RETURN $r \; (mod \; H)$ for all $r \in results$
			\end{algorithmic}
		\end{algorithm}

		This method ``forgets'' the link between generated constraints and the tasks
		(with the associated $A_i$) they come from. Constraints are instead grouped by their
		primal power modulo $p$ and each $p$ is associated with a set of $x_c$ values
		for $c$ in $C_p$, acting similarly as the $A_i$ set of each task. Due to this
		simplification, Algorithm~\ref{alg:multiACRP} might output ``false''
		solutions. This isn't a major drawback as checking whether a solution is
		correct can be done in linear time in the number of tasks (for each task
		$\tau_i$, check whether at least one $a \in A_i$ satisfies $t_d \equiv a \;
		(mod \; T_i)$) and the only interesting result is the minimum $t_d$ which
		means that using a minimum heap structure containing the results we only need
		to check solutions for correctness until the first (hence minimum) correct
		solution is found.

		Both Algorithm~\ref{alg:algoCRP} and Algorithm~\ref{alg:multiACRP} require the
		computation of one modular inversion per constraint and per $a_i$ value, which
		is done in a linear number of modular inversions per task. However, to
		find all DITs Algorithm~\ref{alg:algoCRP} has to be executed once for each possible
		combination of $a_i$ values, which is exponential in the number of tasks.
		Algorithm~\ref{alg:multiACRP} has the same theoretical complexity as the
		computation of all DITs still requires an exponential number of answers in the
		number of tasks, but it is nonetheless significantly faster in practice. The
		running time of Algorithm~\ref{alg:multiACRP} is also less sensitive to the
		size of the $A_i$ sets per task, which means that this method
		remains practical when tasks deadlines are much smaller than periods.
